### 2022-05-09 discussion.tex
Complete content-wise. Waiting on final feedback.

\chapter{Discussion}
%In this section you discuss any issues that came up while developing
%the system.  If you found something particularly interesting,
%difficult, or an important learning experience, put it here.  This is
%also a good place to put additional figures and data.
Mean Absolute Percentage Error (MAPE) does not tell the whole story of a model's performance. The true measure of performance is how useful the model output is to energy producers. The model's confidence in a prediction is almost as important as the accuracy of the prediction and the total energy produced over the day matters.
The forecast downward shortwave irradiance, the largest factor in measured irradiance, rarely drops. It is almost exclusively on very high variance days that it deviates from a sine wave. The downward diffuse shortwave irradiance, the other network input data value, is a much smaller part of the total irradiance and has, to the bare eye, seemingly no correlation with the observation values. Despite this, the models show generally fantastic performance in predicting the irradiance.
The full model was able to utilize the forecast data from the surrounding area to achieve noticeably better performance. Of course, since the limited model uses only one-ninth of the data of the full model, some hyperparameter tuning had to be done to optimize its performance and some of the performance difference could be explained by worse optimization.
With the data at hand, it is not possible to separate the performance of the weather forecast model producing our input data and the performance of our models. We do not know how close this model comes to maximizing the information available in the input data.

Fig.~\ref{fig:disc_low} shows side-by-side the two models' performance for the same calm day. The full model was able to track the observations more closely with more confidence. This was typical for all calm days.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/disc/disc_med_full.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/low/low_g_1.png}}\qquad
    \caption{Comparison of model performance on a clear day.
    \label{fig:disc_low}}
\end{figure}

In Fig.~\ref{fig:disc_med} we can see a comparison between the two models on a less predictable day. Again, the full model shows better performance in tracking, primarily because the limited model misses an end-of-day prediction. There are similar days where the limited model shows slightly better performance.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}}\qquad
    \caption{Comparison of model performance on a day with some drops in irradiance.
    \label{fig:disc_med}}
\end{figure}

Fig.~\ref{fig:disc_high} shows a clear example of the performance difference between the models on erratic days. The full model, while still having a high error gets somewhat close to keeping the observations within its much smaller 90\% confidence interval than the limited model. The limited model also predicts much more irradiance over the day than the, already too high, full model.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_4.png}}\qquad
    \caption{Comparison of model performance on an erratic day.
    \label{fig:disc_high}}
\end{figure}

>>>>

\chapter{Discussion}
%In this section you discuss any issues that came up while developing
%the system.  If you found something particularly interesting,
%difficult, or an important learning experience, put it here.  This is
%also a good place to put additional figures and data.
Mean Absolute Percentage Error (MAPE) does not tell the whole story of a model's performance. The true measure of performance is how useful the model output is to energy producers. The model's confidence in a prediction is almost as important as the accuracy of the prediction, and the total energy produced over the day matters.
The forecast downward shortwave irradiance, the largest factor in measured irradiance, rarely drops. It is almost exclusively on very high variance days that it deviates from a sine wave. The downward diffuse shortwave irradiance, the other network input data value, is a much smaller part of the total irradiance and has, to the bare eye, seemingly no correlation with the observation values. Despite this, the models show generally fantastic performance in predicting the irradiance.
The full model was able to utilize the forecast data from the surrounding area to achieve noticeably better performance. Of course, since the limited model uses only one-ninth of the data of the full model, some hyperparameter tuning had to be done to optimize its performance and some of the performance difference could be explained by worse optimization.
With the data at hand, it is not possible to separate the performance of the weather forecast model producing our input data and the performance of our models. We do not know how close this model comes to maximizing the information available in the input data.

Figure~\ref{fig:disc_low} shows side-by-side the two models' performance for the same calm day. The full model was able to track the observations more closely with more confidence. This was typical for all calm days.\\

\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/disc/disc_med_full.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/low/low_g_1.png}}\qquad
    \caption{Comparison of model performance on a clear day.
    \label{fig:disc_low}}
\end{figure}

In Figure~\ref{fig:disc_med} we can see a comparison between the two models on a less predictable day. Again, the full model shows better performance in tracking, primarily because the limited model misses an end-of-day prediction. There are similar days where the limited model shows slightly better performance.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}}\qquad
    \caption{Comparison of model performance on a day with some drops in irradiance.
    \label{fig:disc_med}}
\end{figure}

Figure~\ref{fig:disc_high} shows a clear example of the performance difference between the models on erratic days. The full model, while still having a high error gets somewhat close to keeping the observations within its much smaller 90\% confidence interval than the limited model. The limited model also predicts much more irradiance over the day than the, already too high, full model.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_4.png}}\qquad
    \caption{Comparison of model performance on an erratic day.
    \label{fig:disc_high}}
\end{figure}

### 2022-05-09 introduction.tex
Complete content-wise. Waiting on final feedback.

Solar power production is highly dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}. A single cloud can adversely impact the production of a solar farm. Such unreliability in a power production system is undesirable since production must match demand \cite{lee_forecasting_2018}. When solar power production falls, the power demand it was serving must be fulfilled by other means. Fossil fueled power stations are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}. These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.\\
Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}. This notice enables planning for the dips in power production and making arrangements for the future. This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.\\

>>>>

Solar power production is highly dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016, jimenez_wrf-solar_2016, rognvaldsson_numerical_2013}. A single cloud can adversely impact the production of a solar power station. Such unreliability in a power production system is undesirable since production must match demand \cite{lee_forecasting_2018}. When solar power production falls, the power demand it was serving must be fulfilled by other means. Fossil fueled power stations are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}. These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.\\
Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}. This notice enables planning for the dips in power production and making arrangements for the future. This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.\\

### 2022-05-09 methods.tex
Complete content-wise. Waiting on final feedback.

\section{Model}
    \subsection{Temporal Fusion Transformer}
    \textcolor{red}{TODO: add more meat}

@@

    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data.\\
    A Gated Residual Network (GRN), shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.\\
    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.\\
    \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
    \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, show a confidence interval for the predictions \cite{lim_temporal_2020}. \\
    Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.

@@

    \subsubsection{Loss function}\label{sec:loss_function}
    \textcolor{red}{TODO: Flesh out and fix}\\
    We use the quantile loss  Temporal Fusion Transformers is quantile loss, which we use.
    A model's prediction is a random value within a probability distribution \cite{wen_multi-horizon_2018}. If we make multiple predictions we can estimate this probability distribution by seeing within what range a selected proportion of the predictions fall \cite{koenker_quantile_2001}. PyTorch Forecasting's quantile loss function takes the Mean Absolute Error (MAE) of three quantiles, 90\%, 50\% and 10\%. This gives a good picture of the model's confidence in a prediction. A smaller quantile range indicated more confidence.

>>>>

\section{Model}
    \subsection{Temporal Fusion Transformer}

@@

    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data, or datasets where it is unknown which parts are of significance.\\
    The Gated Residual Network (GRN) structure, shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}. It provides the model the flexibility to choose the complexity of its input, more complex where needed, and less complex where not.\\
    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}. Some variables may be more important than others, \enquote{TFT is designed to provide instance-wise variable selection through the use of variable selection networks applied to both static covariates and time-dependent covariates.} \cite{lim_temporal_2020} The variable selection mechanism also allows the model to remove noisy inputs that might have a negative effect on performance. \\
    \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
    \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, display a confidence interval for the predictions \cite{lim_temporal_2020}. \\
    Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.

@@

    \subsubsection{Loss function}\label{sec:loss_function}
    The loss function used for Temporal Fusion Transformers is quantile loss, as it's an integral part of generating the confidence intervals of the model \cite{lim_temporal_2020}.
    A model's prediction is a random value within a probability distribution \cite{wen_multi-horizon_2018}. If we make multiple predictions we can estimate this probability distribution by seeing within what range a selected proportion of the predictions fall \cite{koenker_quantile_2001}. PyTorch Forecasting's quantile loss function takes the Mean Absolute Error (MAE) of three quantiles, 90\%, 50\% and 10\%. This gives a good picture of the model's confidence in a prediction. A smaller quantile range indicates more confidence.
